canaper: Categorical analysis of neo- and paleo-endemism in R

### 2022-09-22 ms/ms.Qmd
Update MS

---
title: "canaper: Categorical analysis of neo- and paleo-endemism in R"
bibliography: references.yaml
csl: methods-in-ecology-and-evolution.csl
highlight-style: github
format:
  docx: default
  pdf:
    include-in-header:
      # keep figures from floating away
      text: |
        \usepackage{float}
        \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
---

@@

To demonstrate usage with a real dataset, we reproduced the analysis of @Mishler2014, who conducted CANAPE on Australian *Acacia* using Biodiverse.
Conducting CANAPE on Australian *Acacia* using `canaper` is simple, requiring only two commands:

@@

We also re-ran the *Acacia* analysis using Biodiverse with the same settings as @Mishler2014 (999 replicates of the `rand_structured` null model) and compared the results with those from `canaper`.
Importantly, there is no expectation that results between the two should match exactly, for two reasons.
First, the null model used between the Biodiverse and `canaper` analyses differ (`rand_structured` and `curveball`, respectively).
`rand_structured` is not currently available in R, but we hope to add this to a future version of `canaper`.
Second, the random communities generated in each run will be different, so the exact *p*-values will also be different.
With a sufficiently high number of random communities, significance (e.g., at the $\alpha = 0.5$ level) is expected to converge, but there may be borderline cases that appear significant in some analyses and non-significant in others.

@@

Computations were carried out on an iMac (2019) with 16 GB RAM and a 3GHz, six core processor.
Approximate compute times were 1.3 hr for Biodiverse and `r times$acacia_canape_cpr_1$minutes` min for `canaper` (sequential, i.e., non-parallel mode, for both).
However, when parallel computing was enabled for `canaper` with two cores, compute time dropped to `r times$acacia_canape_cpr$minutes` min (further increases in number of cores did not significantly decrease time).
This demonstrates that `canaper` can efficiently compute moderately sized datasets with a personal computer.

@@

>>>>

---
title: "canaper: Categorical analysis of neo- and paleo-endemism in R"
bibliography: references.yaml
csl: methods-in-ecology-and-evolution.csl
highlight-style: github
crossref: 
  fig-title: Fig.
  tbl-title: Table
format:
  docx: default
  pdf:
    include-in-header:
      # keep figures from floating away
      text: |
        \usepackage{float}
        \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
---

@@

To demonstrate usage with a real dataset, we reproduced the analysis of @Mishler2014, who conducted CANAPE on Australian *Acacia* using Biodiverse.
All `canaper` analyses were run with `canaper` v`r packageVersion("canaper")` in R v`r R.Version()$major`.`r R.Version()$minor`.
Conducting CANAPE on Australian *Acacia* using `canaper` is simple, requiring only two commands:

@@

We also re-ran the *Acacia* analysis using Biodiverse v3.1 with the same settings as @Mishler2014 (999 replicates of the `rand_structured` null model) and compared the results with those from `canaper`.
Importantly, there is no expectation that results between the two should match exactly, for two reasons.
First, the null model used between the Biodiverse and `canaper` analyses differ (`rand_structured` and `curveball`, respectively).
`rand_structured` is not currently available in R, but we hope to add this to a future version of `canaper`.
Second, the random communities generated in each run will be different, so the exact *p*-values will also be different.
With a sufficiently high number of random communities, significance (e.g., at the $\alpha = 0.5$ level) is expected to converge, but there may be borderline cases that appear significant in some analyses and non-significant in others.

@@

Computations were carried out on an MacBook Pro (2019) with 16 GB RAM and a 1.7 GHz, four core processor.
Approximate compute times were 1.3 hr for Biodiverse and `r times$acacia_canape_cpr_1$minutes` min for `canaper` (sequential, i.e., non-parallel mode, for both).
However, when parallel computing was enabled for `canaper` with two cores, compute time dropped to `r times$acacia_canape_cpr$minutes` min (further increases in number of cores did not significantly decrease time).
This demonstrates that `canaper` can efficiently compute moderately sized datasets with a personal laptop computer.

@@

fig <- a + b + plot_layout(ncol = 1) + plot_annotation(tag_levels = "a")

ggsave(plot = fig, file = "ms/figure_1.pdf", width = 100, height = 200, units = "mm")

fig
```
